### 2017-11-25 gcll.tex
add gcll

\documentclass{ctexart}

\begin{document}

{\linespread{1.0}\selectfont
The baseline skip is set to twice
the normal baseline skip.
Pay attention to the \verb|\par|
command at the end. \par}
In comparison, after the
curly brace has been closed,
everything is back to normal.

\quad

This\hspace{1.5cm}is a space
of 1.5 cm.

\quad

x\hspace{\stretch{1}}
x\hspace{\stretch{3}}
x\hspace{\fill}x

\end{document}