### 2017-11-29 test.tex

\documentclass[a4paper,UTF8]{article}
\usepackage{ctex}
\usepackage{caption}
\usepackage{graphicx}
\begin{document}
\begin{figure}[ht]
\begin{minipage}{0.3\linewidth}
\centerline{\includegraphics[width=1\textwidth]{../Picture/gakki.jpg}}
\centerline{伤心图}
\end{minipage}
\quad
\begin{minipage}{0.3\linewidth}
\centerline{\includegraphics[width=1\textwidth]{../Picture/gakki.jpg}}
\centerline{开心图}
\end{minipage}
\quad
\begin{minipage}{0.3\linewidth}
\centerline{\includegraphics[width=1\textwidth]{../Picture/gakki.jpg}}
\centerline{开心图}
\end{minipage}
\caption*{都是表情图}
\end{figure}
\end{document}